In-silico molecular enrichment and clearance of the human intracranial space

### 2025-01-26 paper/main.tex

From T1-weighted MR images of a 26-year-old, healthy male volunteer~\cite{hodneland2019new},
we first automatically segment the brain parenchyma and CSF spaces,
including the ventricular system and SAS, using
Synthseg~\cite{billot2023robust,billot2023synthseg}
(\Cref{fig:results1}A--B). We next manually adjust the segmentation to
accurately represent the connections between CSF spaces (aqueduct,
median aperture, tentorium cerebelli), smoothen, and finally extract
surface representations of the outer (arachnoid) boundary, the pial
membrane and other interfaces. Conforming to these surface and
interface representations, we generate a tetrahedral mesh $\Omega$
representing the complete intracranial volume as the union of the
parenchyma $\Omega_{\rm PAR}$ and CSF spaces $\Omega_{\rm CSF}$, using
fTetWild~\cite{hu2020fast} (\Cref{fig:csf}A--B). The resulting
computational mesh consists of $233\,592$ vertices and $1\,290\,131$
mesh cells, which vary between $0.22$ mm and $8.9$ mm in mesh cell
diameter. The parenchyma has a total volume of $1318$ ml, whereas the
ventricles and the outer SAS contribute $32$ ml and $329$ ml,
respectively, to a total CSF volume of $361$ml, in agreement with
recent estimates~\cite{hladky2024regulation}.

@@

To represent surface networks of periarterial and perivenous spaces,
we use Kiminaro~\cite{william_silversmith_2021_5539913} to separately
skeletonize time-of-flight angiography (ToF) and quantitative
susceptibility mapping (QSM) images from the same human
subject~\cite{hodneland2019new}. This technique yields networks of
one-dimensional curves, each curve $\Lambda^i$ indicating the
centerline of a blood vessel segment and labeled with its lumen radius
$R_1^i$ (\Cref{fig:results1}C, \Cref{fig:pvs_a}A). We also assign an
outer radius $R_2^i > R_1^i$ to each segment $i$, form the annular
cylinder ensheathing $\Lambda^i$ and define the union of these as the
PVS. The periarterial network graph consists of $12\,708$ edges
connected at $12\,562$ inner nodes and with $147$ end nodes, and the
associated domain is denoted by $\Lambda_a$. We identify three of the
end nodes -- corresponding to the two internal carotid arteries (ICAs)
and the basilar artery (BA) -- as root nodes, and designate the other
ends as leaf nodes. The resulting perivenous network graph consists of
$\fixme{N}$ edges connected at $\fixme{N}$ inner nodes and with
$\fixme{N}$ end nodes, and the associated domain is denoted by
$\Lambda_v$. \discuss{Finally, we label the outer surface of the PVSs
  associated with $\Lambda_a$ and $\Lambda_v$, by $\Gamma_a$ and
  $\Gamma_v$ respectively.}

>>>>

From T1-weighted MR images of a 26-year-old, healthy male volunteer~\cite{hodneland2019new},
we first automatically segment the brain parenchyma and CSF spaces,
including the ventricular system and SAS, using
Synthseg~\cite{billot2023robust,billot2023synthseg}
(\Cref{fig:results1}A--B). We next manually adjust the segmentation to
accurately represent the connections and barriers between CSF spaces (aqueduct,
median aperture, tentorium cerebelli), smoothen, and finally extract
surface representations of the outer (arachnoid) boundary, the pial
membrane and other interfaces. Conforming to these surface and
interface representations, we generate a tetrahedral mesh $\Omega$
representing the complete intracranial volume as the union of the
parenchyma $\Omega_{\rm PAR}$ and CSF spaces $\Omega_{\rm CSF}$, using
fTetWild~\cite{hu2020fast} (\Cref{fig:csf}A--B). The resulting
computational mesh consists of $233\,592$ vertices and $1\,290\,131$
mesh cells, which vary between $0.22$ mm and $8.9$ mm in mesh cell
diameter. The parenchyma has a total volume of $1318$ ml, whereas the
ventricles and the outer SAS contribute $32$ ml and $329$ ml,
respectively, to a total CSF volume of $361$ml, in agreement with
recent estimates~\cite{hladky2024regulation}.

@@

To represent surface networks of periarterial and perivenous spaces,
we use Kiminaro~\cite{william_silversmith_2021_5539913} to separately
skeletonize time-of-flight angiography (ToF) and quantitative
susceptibility mapping (QSM) images from the same human
subject~\cite{hodneland2019new}. This technique yields networks of
one-dimensional curves, each curve $\Lambda^i$ indicating the
centerline of a blood vessel segment and labeled with its lumen radius
$R_1^i$ (\Cref{fig:results1}C, \Cref{fig:pvs_a}A). We also assign an
outer radius $R_2^i > R_1^i$ to each segment $i$, form the annular
cylinder ensheathing $\Lambda^i$ and define the union of these as the
PVS. The periarterial network graph consists of $12\,708$ edges
connected at $12\,562$ inner nodes and with $147$ end nodes, and the
associated domain is denoted by $\Lambda_a$. We identify three of the
end nodes -- corresponding to the two internal carotid arteries (ICAs)
and the basilar artery (BA) -- as root nodes, and designate the other
ends as leaf nodes. The resulting perivenous network graph consists of
$24\,829$ edges connected at $23\,881$ inner nodes and with $949$ end nodes, and the associated domain is denoted by $\Lambda_v$. Finally, we label the outer surface of the PVSs associated with $\Lambda_a$ and $\Lambda_v$, by $\Gamma_a$ and $\Gamma_v$ respectively.
